---
title: Influence of Transmission Type on Vehicle Fuel Consumption - A foray into Regression
  Analysis
sansfont: Calibri Light
output:
  pdf_document:
    highlight: pygments
    includes:
      in_header: header2.tex
    keep_tex: no
    number_sections: no
  html_document:
    highlight: pygments
    keep_md: yes
---


```{r "Data", echo=FALSE, cache=TRUE, }
## Loading the dataset
library(datasets)
data(mtcars)

```

#Output Summaries

## Correlation

```{r "CorrelationsTables", echo=FALSE, message=FALSE}
library(xtable)
library(Hmisc)

simpleCor <- cor(mtcars)
corrs <- rcorr(as.matrix(mtcars), type="spearman")
coCorr <- corrs$r
pvCorr <- round(corrs$P, digits = 4)

print(xtable(simpleCor), file="taa.tex", floating=FALSE)
print(xtable(coCorr), file="tbb.tex", floating=FALSE)


```

\begin{table}[ht]
\centering
\subfloat[Table - Simple Correlation]{\label{tab:tab1a}\scalebox{.5}{\input{./taa}}}\quad
\subfloat[Table - Pearson Correlation]{\label{tab:tab1b}\scalebox{.5}{\input{./tbb}}}
\caption{One-to-One Correllations}
\label{tab:tab1}
\end{table}

## Modelling schemas

```{r "Models1", echo=TRUE}
## Method 1
fitAll <- lm(mpg ~ ., data=mtcars)
fitAlla <- update(fitAll, mpg ~ . - qsec) 
fitAllb <- update(fitAlla, mpg ~ . - qsec - carb) 
fitAllc <- update(fitAllb, mpg ~ . - qsec - carb - gear)
fitAlld <- update(fitAllc, mpg ~ . - qsec - carb - gear - drat)
fitAlle <- update(fitAlld, mpg ~ . - qsec - carb - gear - drat - vs)
fitAllf <- update(fitAlle, mpg ~ . - qsec - carb - gear - drat - vs - cyl)
fitAllg <- update(fitAllf, mpg ~ . - qsec - carb - gear - drat - vs - cyl - disp)
fitAllh <- update(fitAllg, mpg ~ . - qsec - carb - gear - drat - vs - cyl - disp - hp)

meth1Anova <- anova(fitAll, fitAlla, fitAllb, fitAllc, fitAlld, fitAlle, fitAllf, fitAllg, fitAllh)

library(xtable)

print(xtable(summary(fitAll)), file="tall.tex", floating=FALSE)
print(xtable(summary(fitAlla)$coef), file="talla.tex", floating=FALSE)
print(xtable(summary(fitAllb)$coef), file="tallb.tex", floating=FALSE)
print(xtable(summary(fitAllc)$coef), file="tallc.tex", floating=FALSE)
print(xtable(summary(fitAlld)$coef), file="talld.tex", floating=FALSE)
print(xtable(summary(fitAlle)$coef), file="talle.tex", floating=FALSE)
print(xtable(summary(fitAllf)$coef), file="tallf.tex", floating=FALSE)
print(xtable(summary(fitAllg)$coef), file="tallg.tex", floating=FALSE)

meth1Anova

```

\begin{table}[ht]
\centering
\subfloat[Table - All]{\label{tab:tab1a}\scalebox{.5}{\input{./tall}}}\quad
\subfloat[Table - a]{\label{tab:tab1b}\scalebox{.5}{\input{./talla}}}\quad
\subfloat[Table - b]{\label{tab:tab1c}\scalebox{.5}{\input{./tallb}}}\quad
\subfloat[Table - c]{\label{tab:tab1d}\scalebox{.5}{\input{./tallc}}}\quad
\subfloat[Table - d]{\label{tab:tab1e}\scalebox{.5}{\input{./talld}}}\quad
\subfloat[Table - e]{\label{tab:tab1f}\scalebox{.5}{\input{./talle}}}\quad
\subfloat[Table - f]{\label{tab:tab1g}\scalebox{.5}{\input{./tallf}}}\quad
\subfloat[Table - g]{\label{tab:tab1g}\scalebox{.5}{\input{./tallg}}}
\caption{Step wise - back}
\label{tab:tab1}
\end{table}


```{r "Models2", echo=TRUE}

## Method 2
fit1 <- lm(mpg ~ am:wt, mtcars)
fit1a <- update(fit1, mpg ~ am:wt + hp)
fit1b <- update(fit1a, mpg ~ am:wt + hp*cyl)
fit1c <- update(fit1b, mpg ~ am:wt + hp*cyl*disp)
fit1d <- update(fit1c, mpg ~ am:wt + hp*cyl*disp*drat)

meth2Anova <- anova(fit1, fit1a, fit1b, fit1c, fit1d)

library(xtable)

print(xtable(summary(fit1)), file="t1.tex", floating=FALSE)
print(xtable(summary(fit1a)$coef), file="t1a.tex", floating=FALSE)
print(xtable(summary(fit1b)$coef), file="t1b.tex", floating=FALSE)
print(xtable(summary(fit1c)$coef), file="t1c.tex", floating=FALSE)
print(xtable(summary(fit1d)$coef), file="t1d.tex", floating=FALSE)

meth2Anova
```



\begin{table}[ht]
\centering
\subfloat[Table - All]{\label{tab:tab2a}\scalebox{.5}{\input{./t1}}}\quad
\subfloat[Table - a]{\label{tab:tab2b}\scalebox{.5}{\input{./t1a}}}\quad
\subfloat[Table - b]{\label{tab:tab2c}\scalebox{.5}{\input{./t1b}}}\quad
\subfloat[Table - c]{\label{tab:tab2d}\scalebox{.5}{\input{./t1c}}}\quad
\subfloat[Table - d]{\label{tab:tab2e}\scalebox{.5}{\input{./t1d}}}
\caption{Interactions}
\label{tab:tab2}
\end{table}

```{r "Models3", echo=TRUE}

## Method 3
fit2 <- lm(mpg ~ am + wt, mtcars)
fit2a <- update(fit2, mpg ~ am + wt + hp)
fit2b <- update(fit2a, mpg ~ am:wt + hp)


meth3Anova <- anova(fit2, fit2a, fit2b)

library(xtable)

print(xtable(summary(fit2)), file="t2.tex", floating=FALSE)
print(xtable(summary(fit2a)$coef), file="t2a.tex", floating=FALSE)
print(xtable(summary(fit2b)$coef), file="t2b.tex", floating=FALSE)


meth3Anova
```




\begin{table}[ht]
\centering
\subfloat[Table - a]{\label{tab:tab3a}\scalebox{.5}{\input{./t2}}}\quad
\subfloat[Table - b]{\label{tab:tab3b}\scalebox{.5}{\input{./t2a}}}\quad
\subfloat[Table - c]{\label{tab:tab3c}\scalebox{.5}{\input{./t2b}}}\quad
\caption{weight}
\label{tab:tab3}
\end{table}




